---
title: "p8105_hw6_jq2293"
author: "JiyueQin"
date: "November 24, 2018"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Problem 1

This problem uses dataset about homicides in 50 large US cities.

## import and tidy

First, import the data from website and do some cleaning.

```{r tidy, message = FALSE}
homicide_data = read_csv("https://github.com/washingtonpost/data-homicides/raw/master/homicide-data.csv") %>% 
  unite(city_state, city, state, sep = ",", remove = FALSE) %>% 
  mutate(solved = ifelse(disposition == "Closed by arrest", TRUE, FALSE)) %>% 
  filter(! city %in% c("Dallas","Phoenix","Kansas City") &
           city_state != "Tulsa,AL") %>% 
  mutate(victim_race = recode_factor(victim_race, White = "white", .default = "non-white"),
         victim_age = as.numeric(victim_age))
```

The initial cleaning steps include:

1. using `unite()` to add a new variable `city_state`.

2. using `mutate()` to add a new variable `solved`.

3. using `filter()` to remove observations in certain cities which lack inforation of victim race.

4. using `mutate()` to modify `victim_race` into categories of white and non-white and convert `victim_age` into numeric variable.

The resulting dataset `homicide_data` has `r nrow(homicide_data)` rows and `r ncol(homicide_data)` columns.

Some key variables are:

- `victim_race`: factor variable, white as reference.

- `city_state`: character varibale, denotes the place of the crime.  

- `lat` and `lon`: numeric variable, denotes the geo location of the crime.

- `solved`: logical variable, `TRUE`denotes the crime is sloved, meaning closed by arrest.

