---
title: "p8105_hw6_jq2293"
author: "JiyueQin"
date: "November 24, 2018"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Problem 1

This problem uses dataset about homicides in 50 large US cities.

## import and tidy

First, import the data from website and do some cleaning.

```{r tidy, message = FALSE}
homicide_data = read_csv("https://github.com/washingtonpost/data-homicides/raw/master/homicide-data.csv") %>% 
  unite(city_state, city, state, sep = ",", remove = FALSE) %>% 
  mutate(solved = as.numeric(disposition == "Closed by arrest")) %>% 
  filter(! city %in% c("Dallas","Phoenix","Kansas City") &
           city_state != "Tulsa,AL") %>% 
  mutate(victim_race = recode_factor(victim_race, White = "white", .default = "non-white"),
         victim_age = as.numeric(victim_age))
```

The initial cleaning steps include:

1. using `unite()` to add a new variable `city_state`.

2. using `mutate()` to add a new variable `solved`.

3. using `filter()` to remove observations in certain cities which lack inforation of victim race.

4. using `mutate()` to modify `victim_race` into categories of white and non-white and convert `victim_age` into numeric variable.

The resulting dataset `homicide_data` has `r nrow(homicide_data)` rows and `r ncol(homicide_data)` columns.

Some key variables are:

- `victim_race`: factor variable, white as reference.

- `city_state`: character varibale, denotes the place of the crime.  

- `lat` and `lon`: numeric variable, denotes the geo location of the crime.

- `solved`: numeric variable(binary), 1 denotes the crime is sloved, meaning closed by arrest. Otherwise, take 0.

## regression in Baltimore

Then, fit a logistic regression to the records in Baltimore. Outcome is resolved vs unresolved and predictors are victim age, sex and race. 

```{r reg}
fit_logit_Bal =
  homicide_data %>% 
  filter(city == "Baltimore") %>% 
  glm(solved ~ victim_age + victim_sex + victim_race, data = ., family = binomial()) 

fit_logit_Bal %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         term = str_replace(term, "sex", "sex:"),
         term = str_replace(term, "race", "race:"))%>%
  select(term, log_OR = estimate, OR, p.value) %>% 
  knitr::kable(digits = 3)
  
result = confint(fit_logit_Bal) %>%
  as.data.frame() %>%
  mutate(term = row.names(.)) %>% 
  full_join(., broom::tidy(fit_logit_Bal), by = "term") %>% 
  select(term, log_OR = estimate, lower_log_OR = "2.5 %", upper_log_OR = "97.5 %") %>% 
  mutate(OR = exp(log_OR),
         lower_OR = exp(lower_log_OR),
         upper_OR = exp(upper_log_OR),
         term = str_replace(term, "sex", "sex:"),
         term = str_replace(term, "race", "race:"))

 knitr::kable(result, digits = 3)
```

From the results of the regression, we can see age, sex and race are all significant predictors at 0.05 level.

For solving homicides comparing non-white victims to white victims keeping all other variables fixed, the estimated odds ratio is `r round(result[4, 5], 3)`, the 95% confidence is (`r round(result[4, 6], 3)`, `r round(result[4, 7], 3)`).

This indicates non-white victims are more likely to have unsolved homicides than white victims. 

